\chapter*{ \begin{center} 
    \textbf{\LARGE ABSTRACT}  % Large, bold, and centered heading  
\end{center} }  
\addcontentsline{toc}{chapter}{ABSTRACT}  
\vspace{-2cm}  % Adjust this value if needed to reduce space further  
 

The rapid growth of digital transportation services has significantly changed the way people book and manage travel. Online bus booking systems have emerged as an efficient solution for passengers to search routes, reserve seats, and make payments conveniently.
\\\\
This project titled “Bus Ticket Booking System” focuses on the design and development of a frontend-based web application that enables users to search buses, select seats, register and log in, and proceed to checkout securely. The system also includes an admin interface for managing buses, routes, schedules, and fares.
\\\\
The frontend of the system is developed using modern web technologies such as HTML, CSS, JavaScript, and React.js, with emphasis on usability, responsiveness, and modular design. User authentication, seat selection, passenger details entry, and payment workflow are simulated using frontend logic and browser storage.
\\\\
An admin module is included to demonstrate administrative privileges such as adding, editing, and deleting bus details, modifying routes, timings, and fare structures. Backend services are assumed and represented using mock data, as backend integration is outside the scope of the current academic phase.
\\\\
This project demonstrates how a well-designed frontend can enhance user experience, improve accessibility, and provide a strong foundation for a complete full-stack online bus booking system.
